A new method to estimate dark matter halo concentrations

### 2015-09-10 paper/article.tex
minor edits and some new paragraphs

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2},
\label{eq:definition}
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is known as the scale radius, the radius that marks the transition
between the power law scaling $\rho\propto r^{-1}$ for
$r<r_s$ and $\rho\propto r^{-3}$ for  $r>r_s$.

@@

The method we use to generate the halos is based on the integrated
mass profile.
We start by fixing the desired concentration $c$ and total number of
particles $N$ in the mock halo
With this values we define the mass element as $\delta m = 1/M$, corresponding
to the mass of each particle such that the total mass is one.
Then for each particle $i=1,\ldots,N$, we find the value of $r_i$ such that
the difference
%
\begin{equation}
m(<r_i;c) - i \cdot \delta m
\end{equation}
%
is zero using Ridders' method.

@@

\begin{figure*}

  \includegraphics[width=0.45\textwidth]{mock_percentual_diff_conc_20.pdf}
  \includegraphics[width=0.45\textwidth]{mock_percentual_diff_conc_20000.pdf}

\caption{Cumulative distribution of the fractional difference, $D$, between
  the input concentration in the mock halo generator, $c_{in}$ and the
  measurement by our MCMC code, $c_{out}$. Each curve corresponds to
  halos generated with an input concentration in a different range for
  a number of particles $N=20$ (right panel) and $N=20000$ (left panel).
    \label{fig:results_mocks_conc_20}}
\end{figure*}

{\bf Hace falta discutir las figuras.}

\subsubsection{The impact of different methods}

@@

We use data from the MultiDark cosmological simulation.
It follows the non-linear evolution of a dark matter density field
sampled with $2048^3$ particles over a cubic box of $1000$ \hMpc on a side.
The data is publicly available through \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.

We build a sample of all halos located in a cubic sub-volume of $100$
\hMpc on a side centered on the most massive halo in the simulation at
$z=0$ which corresponds to the \texttt{miniMDR1} tables in the
database.
From this sample we select all the halos at $z=0$ detected with a
Friends-of-Friends (FoF) algorithm with masses in the interval
$10^{11}\leq M_{\rm FoF}/\hMsun \leq 10^{15}$.
The FoF algorithm used with a linking length of $0.17$ times the average
inter-particle distance. This choice translates into an overdensity
$\Delta_h\sim 400-700$ dependent on the halo concentration
\citep{More2011}.
Finally, for each FoF halo we select all the particles that
belong to it.

From this set of particles we follow the procedure spelled out in
Section \ref{sec:method} with $\Delta_h=740$  (corresponding to $200$
times the critical density) to find the halo concentration.
This choice makes that our overdensities are fully included inside the
original FoF particle group.
We only report results from overdensities with at least $100$ particles.
Finally, we store the values obtained for the virial radius, virial
mass and concentration.

Figure \ref{fig:mdv} shows the results that compare the concentration
values in the simulated halos from the three different methods.
At zeroth order the results are consistent.
However, the concentration values from the velocity method produces
concentrations that are $XXX$ times larger on average than the results
from our method.
On the other hand, the concentrations from the velocity method are
$XXX$ times lower than in our method.

%Running another test with halos whose concentrations vary as a normal
%distribution of variance ${\sigma}^2$ we found that the concentrations
%obtained by the method of circular velocity are a 35\% greater than
%the ones obtained by the density method. However the concentrations
%obtained by the velocity method are barey higher (around 2\% or 3\%)
%than the ones obtained by our method. We did not find any significant
%variation in the concentration values obtained by any method by
%changing ${\sigma}^2$ in contrast with FIXME.

Figure \ref{fig:concentration} shows the mass-concentration
relationship from the N-body data with the results of the three
different methods.
The central lines correspond the median and the
shadowed region indicates the quartiles.
In this relationship our methods represents a compromise between the
velocity and density methods.
The slope is flatter and the overall normalization is below the
velocity method and above the density.
We also note that at high masses ($M>10^{14}\hMsun$) our method
and the velocity method show a reasonable agreement within the
uncertainties.

>>>>

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2},
\label{eq:definition}
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is the scale radius. This radius marks the transition
between the power law scaling $\rho\propto r^{-1}$ for
$r<r_s$ and $\rho\propto r^{-3}$ for  $r>r_s$.

@@

The method we use to generate the halos is based on the integrated
mass profile.
We start by fixing the desired concentration $c$ and total number of
particles $N$ in the mock halo.
With these values we define the mass element as $\delta m = 1/M$, corresponding
to the mass of each particle such that the total mass is one.
Then for each particle, $i=1,\ldots,N$, we find the value of $r_i$ such that
the difference
%
\begin{equation}
m(<r_i;c) - i \cdot \delta m
\end{equation}
%
is zero using Ridders' method.

@@

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{mock_percentual_diff_conc_20.pdf}
  \includegraphics[width=0.45\textwidth]{mock_percentual_diff_conc_20000.pdf}
\caption{Cumulative distribution of the fractional difference, $D$, between
  the input concentration in the mock halo generator, $c_{in}$ and the
  measurement by our MCMC code, $c_{out}$. Each curve corresponds to
  halos generated with an input concentration in a different range for
  a number of particles $N=20$ (right panel) and $N=20000$ (left panel).
    \label{fig:results_mocks_conc}}
\end{figure*}

Figure \ref{fig:results_mocks_conc} shows the integrared distribution
for $D$ for the fits using our method, split into different input
concentrations for two different particle numbers describing the mock
halos. 
On the left we have the case of halos sampled with $N=20$ particles
and on the right we have show results for halos with $N=20000$
particles. 
The scale for $D$ varies by almost three orders of magnitude.
The accuracy of our method increases with larger values of $N$. 
This was already seen in the previous section. 
We explore the exact dependence of $D$ with the number of particles in
the next section when we compare the accuracy of different methods.

There are two interesting in Figure \ref{fig:results_mock_conc}. The
first is that our method tends to underestimate the concentration
and the second is that the degree of this underestimation increases
with the input halo concentration.

\subsubsection{The accuracy of different methods}

@@

We use data from a N-body cosmological simulation that follows the non
non-linear evolution of a dark matter density field sampled with
$512^3$ particles over a cubic box of $400$ \hMpc on a side. 
This simulation was run with the public available version of the
Gadget-2 code. 
The simulation uses a standard $\Lambda$CDM cosmology with values of
$\Omega_m=0.258$, $\Omega_\Lambda=0.742$ and $h=0.72$ for the matter
density parameter, cosmological constant parameter and the reduce
Hubble constant, respectively. 
With these parameters the mass of an individual simulation particle
corresponds to $3.41\times 10^{10}$\hMsun.

{\bf descripcion del halo finder}

{\bf descripcionde separacion por grado de relajacion}.

{\bf describcion de los resultados}
Figure \ref{fig:mdv} shows the results that compare the concentration
values in the simulated halos from the three different methods.
